Experience Report: Distributed Computing for the Common Man

### 2010-03-22 doc/icfp-2010/main.tex
*** empty log message ***

n-queens, SMT solvers, etc.

>>>>

In this section, we show the potential of our library with two 
computationally-extensive examples.

\subsection{N-queens}

The first example is the classical $N$-queens problem, where we
compute the total number of solutions for a given $N$.
The following table shows execution timings for various values of $N$
and our three different implementations modules \of{Sequential}, \of{Cores} and
\of{Network}. The purpose of this experiment is to measure the speedup
ratio w.r.t. the sequential implementation. Therefore, all
computations are performed on the same machine, a 8 cores Intel Xeon
3.2 GHz running Debian Linux. The sequential implementation uses a
single core. The cores implementation uses the 8 cores of
the machine. The network implementation uses 8 workers running locally and
a master running on a remote machine (incurs communication cost).

The first column shows the value of $N$.
Each computation is divided into a certain number of tasks,
which is shown in second column.
Then the last three column show execution timings in seconds for the
three implementations. The figures within brackets show the speedup
ratio w.r.t. sequential implementation.
\begin{center}
  \begin{tabular}{|r|r|r|r|r|}
    \hline
    N & \#tasks         & sequential& cores         & network 
    \\\hline\hline
    16 &   16    &  15.2     &   2.04 (7.45$\times$) &  2.35  (6.47$\times$) 
    \\\hline
       &  210    &  15.2     &   2.01 (7.56$\times$) & 21.80  (0.69$\times$)
    \\\hline
    17 &   17    & 107.0     &  17.20 (6.22$\times$) & 16.20  (6.60$\times$)
    \\\hline
       &  240    & 107.0     &  14.00                & 24.90    
    \\\hline
    18 &   18    &           & 123.00               & 125.00     
    \\\hline
       &  272    &           & 103.00               & 124.00
    \\\hline
    19 &   19    &           &                      & 940.00
    \\\hline
       &  306    &           &                      & 819.00      \\\hline
  \end{tabular}
\end{center}
From the table above, it is clear that the \of{Cores} implementation
provides a significant speedup. When the number of tasks is large
enough, the speedup ratio is almost 8, which is also the number of
cores we use. 
% (When the number of tasks is small and not a multiple of
% the number of cores,  for example as in third row with 17 tasks
% executing on 8 cores, the

% TODO comment on network ratios

Where computation cost for each tasks is too small, then communication
cost dominates and could even slow down the execution
w.r.t. sequential implementation, as evident from second row.

\subsection{SMT Solvers}

Here we demonstrate the potential of our library for our application
needs as mentioned in the introduction. Our computing
infrastructure consists of 3 machines with 4, 8 and 8 cores
respectively. The benchmark we use for
experimentation consists of 80 challenging verification conditions
(VC) obtained from the Why platform~\cite{filliatre07cav}.  Each
VC is stored in a file, which is accessible over
NFS. The purpose of the experiment is to check the validity of each VC
using several automated provers (namely Alt-Ergo, Simplify and Z3),
which are all installed on each of the machines in use.

The master program proceeds by reading the 80 file names, turning them
into tasks by multiplying them by the number of provers (240 tasks in total).
Each worker in turn invokes the given prover on the given file.
We set a timeout limit of 600 seconds.
Each task completes with one of the four possible outcomes: \emph{valid},
\emph{invalid}, \emph{cannot-decide} (depending on
whether the VC is valid, invalid or undecided by the prover)
and \emph{timeout}.

\begin{center}
  \begin{tabular}{|r|r|r|r|r|}
    \hline
    prover   & valid & invalid & cannot & timeout 
    \\\hline\hline
    Alt-ergo & 
    \\\hline
    Simplify & 
    \\\hline
    Z3       & 
    \\\hline
  \end{tabular}
\end{center}

% TODO compare total amount of time wrt sequential time
